---
title: "Exploring the BRFSS data"
format: 
  html: 
    embed-resources: true
    fig_height: 4
    highlight: pygments
    theme: spacelab
editor_options: 
  chunk_output_type: console
---

## Setup

### Load packages

```{r load-packages, message = FALSE}
#| label: data1
#| code-fold: true
#| code-summary: load packages
#| results: hide
#| warning: false



library(ggplot2)
library(dplyr)
library(stats)
library(psycho)
library(corrplot)
library(patchwork)

```

### Load data



```{r load-data}
#| label: data2
#| code-fold: true
#| code-summary: Getting data
#| results: hide
#| warning: false


file.exists("/Users/muaro/Documents/GitHub/probability-intro/analysis/week4/brfss2013")
load(file = '/Users/muaro/Documents/GitHub/probability-intro/analysis/week4/brfss2013')
getwd()


```

------------------------------------------------------------------------

## Part 1: Data


The data is Brfss2013


------------------------------------------------------------------------

## Part 2: Research questions

**Research quesion 1:**
What is the relationship between regular exercise (exerany2) and overall life satisfaction (lsatisfy), and general health (genhelth).

**Research quesion 2:**
Are richer people happier?

People with higher level of income (income2) have their emotional needs fulfilled better. (emtsuprt)

**Research quesion 3:**
Does mental health (qlmentl2, qlstres2, qlhlth2) of men and women (sex) react differently to lack of sleep (sleptim1).


------------------------------------------------------------------------




## Question 1 - Happy people exercise

What is the relationship between regular exercise (exerany2) and overall life satisfaction (lsatisfy), and general health (genhelth).


We hypothesis that people who exercise regularly have higher life satisfacton and better health


```{r}
#| label: r1-1
#| code-fold: true
#| code-summary: Code for Research 1-1
#| results: markup
#| warning: false

ds1 <- brfss2013 %>% 
  select(exerany2, lsatisfy, genhlth) 

# tableone::CreateTableOne(data = ds1)
# ds1 %>% count(exerany2)

t1 <- 
  ds1 %>%
  filter(
    !is.na(exerany2),
    !is.na(lsatisfy)
  ) %>% 
  group_by( exerany2, lsatisfy) %>% 
  summarise(
    n_count = n()
  ) %>% 
  group_by(exerany2) %>% 
  mutate(
   total = sum(n_count)
   , prop = n_count/total
  ) %>% 
  ungroup() 
t1 %>% ggplot(aes(x = lsatisfy, y = prop, fill = exerany2)) +
  geom_col(position = position_dodge2(preserve = "single", padding = ), alpha = 0.5) +
  geom_text(aes(label = scales::percent(prop), color = exerany2))

```


## Question 2 - Does money help you?

Are there any depended between the outcome and emotional health?

Which class of people have the most stable emotional health.
```{r}
#| label: r2-1
#| code-fold: true
#| code-summary: Code for Research 2-1
#| results: markup
#| warning: false

ds2 <- brfss2013 %>% select(income2, emtsuprt)
# tableone::CreateTableOne(data = ds2)

# ds2 %>% count(income2)

t2 <- 
  ds2 %>% 
  filter(
    !is.na(income2)
    , !is.na(emtsuprt)
  ) %>% 
  group_by(emtsuprt, income2) %>% 
  summarise(
    num = n()
  ) %>% ungroup() %>% 
  mutate(
    total = sum(num)
    , prop = num/total
  ) %>% ungroup() 
  # print(n = Inf)
# t2


t2 %>% 
  ggplot(aes(x = emtsuprt,y = prop, fill = emtsuprt)) + 
  geom_col(position = position_dodge(width = 2), width = 0.8) +
  facet_wrap(~ income2) +
  #geom_text(aes(label =  scales::percent("Always")), vjust = -3) +
    labs(
      title = "The distribution of people with different outcomes"
      ,x = "Frequensy of needed help"
      ,y = "The percentage of people"
      , subtitle = "The distribution of people's needed of emotional health with total number of people"
    ) +
  theme_minimal() + 
   theme(axis.text.x = element_blank())

ds2 %>% 
  filter(
    !is.na(income2)
    , !is.na(emtsuprt)
  ) %>% 
  group_by(emtsuprt, income2) %>% 
  summarise(
    num = n()
  ) %>% ungroup() %>% 
  group_by(income2) %>% 
  mutate(
    total = sum(num)
    , prop = num/total
  ) %>% ungroup() %>% ggplot(aes(x = emtsuprt,y = prop, fill = emtsuprt)) + 
  geom_col(position = position_dodge(width = 2), width = 0.8) +
  facet_wrap(~ income2) +
  #geom_text(aes(label =  scales::percent("Always")), vjust = -3) +
    labs(
      title = "The distribution of people with different outcomes"
      ,x = "Frequensy of needed help"
      ,y = "The percentage of people"
      , subtitle = "The distribution of people's needed of emotional health in each outcome group"
      , fill = "Frequency"
    ) +
  theme_minimal() + 
   theme(axis.text.x = element_blank())

```

Outcome does not affect on emotional health.


## Question 3 - Is lack of sleep serious proplem?  



How different men's and women's time of sleep? 

```{r}
#| label: r3-1
#| code-fold: true
#| code-summary: Code for Research 3-1
#| results: markup
#| warning: false

ds3 <- brfss2013 %>% 
  select(qlmentl2, qlstres2, qlhlth2, sex, sleptim1)

# tableone::CreateTableOne(data = ds3)

t31 <- 
  ds3 %>% 
  filter(
    !is.na(sex)
    , !is.na(sleptim1)
    , sleptim1 < 15
  ) %>% 
  group_by(sleptim1,sex) %>% 
  summarise(
    num = n()
  ) %>% ungroup() %>% 
  mutate(
    total = sum(num)
  ) %>% 
  group_by(sex) %>% 
  mutate(
    total_s = sum(num)
    , prop_s = num/total_s
    , mean = weighted.mean(sleptim1, w = num)
  )

mean_female <- round(t31$mean[1], 2) 
   

mean_male <-  round(t31$mean[2], 2)


t31 %>% 
  filter(
    sleptim1 < 13
  ) %>% 
  ggplot(aes(x = sleptim1, y = prop_s, fill = sex)) + 
  geom_col(position = position_dodge2(preserve = "single", padding = )) +
  scale_fill_manual(values = c("Male" = "royalblue1", "Female" = "violet")) + 
  geom_vline(xintercept = mean_male, color = "royalblue4", linetype = "dashed", linewidth = 1) +
  geom_vline(xintercept = mean_female, color = "violetred1", linetype = "dashed", linewidth = 1) +
  theme_minimal() +
  labs(
    title = "Time of sleep (men and women)"
    , x = "Sleep time"
    , y = "Percentage of people"
    , fill = "Sex"
  )
  
  

```

**How time of sleep affect to depression between men and female? **


```{r}
#| label: r3-2
#| code-fold: true
#| code-summary: Code for Research 3-2
#| results: markup
#| warning: false

t321 <- ds3 %>% select(qlmentl2, sex, sleptim1) %>% 
  group_by(qlmentl2, sleptim1, sex) %>% 
  summarise(
    num = n()
  ) %>% ungroup() %>% 
  filter(
    !is.na(qlmentl2)
    , !is.na(sleptim1)
    , sex == "Male"
  ) %>% group_by(qlmentl2) %>% 
  mutate(
    mean_sleep = round(mean(sleptim1), 1)
  ) %>% ungroup() %>% distinct() %>% 
  select(qlmentl2, num, mean_sleep) %>% 
  group_by(qlmentl2, mean_sleep) %>% 
  summarise(
    num = sum(num)
  ) %>% 
  ungroup() %>%
  mutate(
    total = sum(num)
    , prop = round(num/total, 2)
  ) 
  
g31 <- t321 %>% ggplot(aes(x = qlmentl2, y = prop)) +
  geom_col() +
  geom_text(aes(label = mean_sleep), vjust = -1, size = 3)+
  scale_y_continuous(limits = c(0,0.8), breaks = seq(0,0.8,0.1))+
    scale_x_continuous(breaks = seq(0,30,5)) +
  annotate("text", x = mean(range(t321$qlmentl2)), y = 0.4, label = round(mean(t321$mean_sleep),1), size = 5, color = "darkgreen") +
  labs(
    title = "For men"
   , x = "Days with depression (Last 30 days)"
   , y = "Percantage of men"

  ) +
  theme_minimal()


t322 <- ds3 %>% select(qlmentl2, sex, sleptim1) %>% 
  group_by(qlmentl2, sleptim1, sex) %>% 
  summarise(
    num = n()
  ) %>% ungroup() %>% 
  filter(
    !is.na(qlmentl2)
    , !is.na(sleptim1)
    , sex == "Female"
  ) %>% group_by(qlmentl2) %>% 
  mutate(
    mean_sleep = round(mean(sleptim1), 1)
  ) %>% ungroup() %>% distinct() %>% 
  select(qlmentl2, num, mean_sleep) %>% 
  group_by(qlmentl2, mean_sleep) %>% 
  summarise(
    num = sum(num)
  ) %>% 
  ungroup() %>%
  mutate(
    total = sum(num)
    , prop = round(num/total, 2)
  ) 

g32 <- t322 %>% 
  filter(
    qlmentl2 != 29
  ) %>% ggplot(aes(x = qlmentl2, y = prop)) +
  geom_col() +
  geom_text(aes(label = mean_sleep), vjust = -1, size = 3) +
  scale_y_continuous(limits = c(0,0.8), breaks = seq(0,0.8, 0.1 )) +
  scale_x_continuous(breaks = seq(0,30,5)) +
  annotate("text", x = mean(range(t322$qlmentl2)), y = 0.4, label = round(mean(t322$mean_sleep),1), size = 5, color = "darkgreen") +
  labs(
    title = "For women"
    , x = "Days with depression (Last 30 days)"
    , y = "Percantage of women"

  ) +
  theme_minimal()

g32 + g31

```

**How sleep time affect of number of fill of energy days?**

```{r}
#| label: r3-3
#| code-fold: true
#| code-summary: Code for Research 3-3
#| results: markup
#| warning: false


t331 <- ds3 %>% select(qlhlth2, sex, sleptim1) %>% 
  group_by(qlhlth2, sleptim1, sex) %>% 
  summarise(
    num = n()
  ) %>% ungroup() %>% 
  filter(
    !is.na(qlhlth2)
    , !is.na(sleptim1)
    , sex == "Male"
  ) %>% group_by(qlhlth2) %>% 
  mutate(
    mean_sleep = round(mean(sleptim1), 1)
  ) %>% ungroup() %>% distinct() %>% 
  select(qlhlth2, num, mean_sleep) %>% 
  group_by(qlhlth2, mean_sleep) %>% 
  summarise(
    num = sum(num)
  ) %>% 
  ungroup() %>%
  mutate(
    total = sum(num)
    , prop = round(num/total, 2)
  ) 
  
g33 <- t331 %>% ggplot(aes(x = qlhlth2, y = prop)) +
  geom_col() +
  geom_text(aes(label = mean_sleep), vjust = -1, size = 2)+
  scale_y_continuous(limits = c(0,0.5), breaks = seq(0,0.5,0.1))+
    scale_x_continuous(breaks = seq(0,30,5)) +
  annotate("text", x = mean(range(t331$qlhlth2)), y = 0.4, label = round(mean(t331$mean_sleep),1), size = 5, color = "darkgreen") +
  labs(
    title = "For men"
    , x = "Days Full Of Energy (last 30)"
    , y = "Percentage of men"
  ) +
  theme_minimal()


t332 <- ds3 %>% select(qlhlth2, sex, sleptim1) %>% 
  group_by(qlhlth2, sleptim1, sex) %>% 
  summarise(
    num = n()
  ) %>% ungroup() %>% 
  filter(
    !is.na(qlhlth2)
    , !is.na(sleptim1)
    , sex == "Female"
  ) %>% group_by(qlhlth2) %>% 
  mutate(
    mean_sleep = round(mean(sleptim1), 1)
  ) %>% ungroup() %>% distinct() %>% 
  select(qlhlth2, num, mean_sleep) %>% 
  group_by(qlhlth2, mean_sleep) %>% 
  summarise(
    num = sum(num)
  ) %>% 
  ungroup() %>%
  mutate(
    total = sum(num)
    , prop = round(num/total, 2)
  ) 

g34 <- t332 %>% 
   ggplot(aes(x = qlhlth2, y = prop)) +
  geom_col() +
  geom_text(aes(label = mean_sleep), vjust = -1, size = 2) +
  scale_y_continuous(limits = c(0,0.5), breaks = seq(0,0.5, 0.1 )) +
  scale_x_continuous(breaks = seq(0,30,5)) +
  annotate("text", x = mean(range(t332$qlhlth2)), y = 0.4, label = round(mean(t332$mean_sleep),1), size = 5, color = "darkgreen") +
  labs(
    title = "For women"
    , x = "Days Full Of Energy (last 30)"
    , y = "Percentage of women"
  ) +
  theme_minimal()

g33 + g34


```

How sleep time affect to number of days when you felt anxious? 

```{r}
#| label: r3-4
#| code-fold: true
#| code-summary: Code for Research 3-4
#| results: markup
#| warning: false

t331 <- ds3 %>% select(qlstres2, sex, sleptim1) %>% 
  group_by(qlstres2, sleptim1, sex) %>% 
  summarise(
    num = n()
  ) %>% ungroup() %>% 
  filter(
    !is.na(qlstres2)
    , !is.na(sleptim1)
    , sex == "Male"
  ) %>% group_by(qlstres2) %>% 
  mutate(
    mean_sleep = round(mean(sleptim1), 1)
  ) %>% ungroup() %>% distinct() %>% 
  select(qlstres2, num, mean_sleep) %>% 
  group_by(qlstres2, mean_sleep) %>% 
  summarise(
    num = sum(num)
  ) %>% 
  ungroup() %>%
  mutate(
    total = sum(num)
    , prop = round(num/total, 2)
  ) 
  
g33 <- t331 %>% ggplot(aes(x = qlstres2, y = prop)) +
  geom_col() +
  geom_text(aes(label = mean_sleep), vjust = -1, size = 2)+
  scale_y_continuous(limits = c(0,0.5), breaks = seq(0,0.5,0.1))+
    scale_x_continuous(breaks = seq(0,30,5)) +
  annotate("text", x = mean(range(t331$qlstres2)), y = 0.4, label = round(mean(t331$mean_sleep),1), size = 5, color = "darkgreen") +
  labs(
    title = "For men"
    , x = "Days felt anxious"
    , y = "Percentage of men"
  ) +
  theme_minimal()


t332 <- ds3 %>% select(qlstres2, sex, sleptim1) %>% 
  group_by(qlstres2, sleptim1, sex) %>% 
  summarise(
    num = n()
  ) %>% ungroup() %>% 
  filter(
    !is.na(qlstres2)
    , !is.na(sleptim1)
    , sex == "Female"
  ) %>% group_by(qlstres2) %>% 
  mutate(
    mean_sleep = round(mean(sleptim1), 1)
  ) %>% ungroup() %>% distinct() %>% 
  select(qlstres2, num, mean_sleep) %>% 
  group_by(qlstres2, mean_sleep) %>% 
  summarise(
    num = sum(num)
  ) %>% 
  ungroup() %>%
  mutate(
    total = sum(num)
    , prop = round(num/total, 2)
  ) 

g34 <- t332  %>% ggplot(aes(x = qlstres2, y = prop)) +
  geom_col() +
  geom_text(aes(label = mean_sleep), vjust = -1, size = 2) +
  scale_y_continuous(limits = c(0,0.5), breaks = seq(0,0.5, 0.1 )) +
  scale_x_continuous(breaks = seq(0,30,5)) +
  annotate("text", x = mean(range(t332$qlstres2)), y = 0.4, label = round(mean(t332$mean_sleep),1), size = 5, color = "darkgreen") +
  labs(
    title = "For women"
    , x = "Days felt anxious"
    , y = "Percentage of women"
  ) +
  theme_minimal()

g33 + g34
```

