---
title: "Exploring the BRFSS data"
output: 
  html_document: 
    fig_height: 4
    highlight: pygments
    theme: spacelab
editor_options: 
  chunk_output_type: console
---

## Setup

### Load packages

```{r load-packages, message = FALSE}
library(ggplot2)
library(dplyr)
library(stats)
library(psycho)
library(corrplot)
library(patchwork)

```

### Load data



```{r load-data}
load("./data-private/raw/intro-prob/brfss2013")
ds0 <- brfss2013
```

------------------------------------------------------------------------

## Part 1: Data

The data is Brfss2013
------------------------------------------------------------------------

```{r}
ds0 %>% count(dispcode)

ds0 %>% dim() # 330 columns, too many to display
ds0[,1:10] %>% glimpse() 
# rely on the supplied codebook to interpret the columns
# What is the row in this data set? What's the grain?
ds0 %>% 
  # filter(dispcode == "Completed interview") %>% 
  summarize(
    row_count = n()
    ,id_count = n_distinct(seqno, X_psu,idate)
  )

ds0 %>% select(seqno) %>% explore::describe_all()

ds0 %>% group_by(seqno) %>% count() %>% 
  filter(n>1)

ds0 %>% filter(seqno==2013000001) %>% View()
```


## Part 2: Research questions

**Research quesion 1:**
What is the relationship between regular exercise (exerany2) and overall life satisfaction (lsatisfy), and general health (genhelth).

**Research quesion 2:**
Are richer people happier?

People with higher level of income (income2) have their emotional needs fulfilled better. 

**Research quesion 3:**
Does mental health (qlmentl2, qlstres2, qlhlth2) of men and women (sex) react differently to lack of sleep (sleptim1).


------------------------------------------------------------------------

## Part 3: Exploratory data analysis

NOTE: Insert code chunks as needed by clicking on the "Insert a new code chunk" button (green button with orange arrow) above. Make sure that your code is visible in the project you submit. Delete this note when before you submit your work.

```{r}

```


## Question 1 - Happy people exercise

> RESEARCH QUESTION 1 - What is the relationship between regular exercise (exerany2), overall life satisfaction (lsatisfy), and general health (genhelth).


We hypothesize that people who exercise regularly have higher life satisfaction and better health


```{r}
#| label: r1-1
#| code-fold: true
#| code-summary: Code for Research 1-1
#| results: markup
#| warning: false

# let's create a data set that would help us focus and explore the variables of interest.
ds1 <- 
  brfss2013 %>% 
  select(exerany2, lsatisfy, genhlth) 
ds1 %>% labelled::look_for()
tableone::CreateTableOne(data = ds1)
# 
#   mutate(across(where(is.factor), ~ forcats::fct_explicit_na(.x, na_level = "NA")))


ds1 %>% count(exerany2)

ds1 %>% labelled::look_for()
table(ds1$exerany2, ds1$lsatisfy)

source("./analysis/week4/binary-categorical-functions.R")
ds1 %>% make_bi_freq_table("exerany2")
ds1 %>% make_bi_freq_table("exerany2","lsatisfy")

ds1 %>% make_bi_freq_graph("exerany2",)
ds1 %>% run_contingency("exerany2","lsatisfy")


t1 <- 
  ds1 %>%
  filter(
    !is.na(exerany2),
    !is.na(lsatisfy)
  ) %>% 
  group_by( exerany2, lsatisfy) %>% 
  summarise(
    n_count = n()
  ) %>% 
  group_by(exerany2) %>% 
  mutate(
   total = sum(n_count)
   , prop = n_count/total
  ) %>% 
  ungroup() 
t1 %>% ggplot(aes(x = lsatisfy, y = prop, fill = exerany2)) +
  geom_col(position = position_dodge2(preserve = "single", padding = ), alpha = 0.5) +
  geom_text(aes(label = scales::percent(prop), color = exerany2))

t1
```


**Research quesion 2:**

```{r}

```

**Research quesion 3:**

```{r}

```
