---
title: "Quiz: Sampling Distributions"
output: learnr::tutorial
runtime: shiny_prerendered
tutorial:
  id: sampling-distribution-quiz
  name: Sampling Distribution Quiz
---

```{r setup, include=FALSE}
library(learnr)
knitr::opts_chunk$set(echo = FALSE)
```

# Sampling Distributions Quiz

This quiz will test your understanding of sampling distributions and related concepts.

## Question 1: What is a sampling distribution?

```{r question-1, echo=FALSE}
question("What is a sampling distribution?",
  answer("A histogram of values in the population"),
  answer("A distribution of repeated sample means from the same population", correct = TRUE),
  answer("A single sample taken from the population"),
  answer("The distribution of standard deviations across different datasets"),
  correct = "A sampling distribution is formed by repeatedly taking samples of the same size and calculating a statistic like the mean.",
  allow_retry = TRUE
)

```

## Question 2: According to the Central Limit Theorem (CLT), what will the sampling distribution of the sample mean be?

```{r question-2, echo=FALSE}
question("According to the Central Limit Theorem (CLT), the sampling distribution of the sample mean will be approximately normal if:",
  answer("The population is normal"),
  answer("The sample size is small"),
  answer("The population is skewed but the sample size is large", correct = TRUE),
  answer("The samples are taken with replacement"),
  correct = "With large enough sample size, the sampling distribution of the mean is approximately normal regardless of population shape.",
  allow_retry = TRUE
)

```

## Question 3: As the sample size increases, what happens to the standard error (SE) of the sample mean?

```{r question-3, echo=FALSE}
question("As the sample size *n* increases, what happens to the standard error (SE) of the sample mean?",
  answer("It stays the same"),
  answer("It increases"),
  answer("It decreases", correct = TRUE),
  answer("It becomes equal to the population standard deviation"),
  correct = "The SE = population SD / sqrt(n), so increasing n reduces SE.",
  allow_retry = TRUE
)

```

## Question 4: If the original population is heavily skewed, what happens to the distribution of the sample mean?

```{r question-4, echo=FALSE}
question("If the original population is heavily skewed, what happens to the distribution of the sample mean as you increase the sample size?",
  answer("It stays skewed"),
  answer("It becomes more normal", correct = TRUE),
  answer("It becomes bimodal"),
  answer("It becomes uniform"),
  correct = "The CLT shows that the sample mean's distribution becomes more normal as n increases, even if the original population is skewed.",
  allow_retry = TRUE
)
 
```

## Question 5: Why do we simulate multiple samples when studying sampling distributions?

```{r question-5, echo=FALSE}
question("Why do we simulate multiple samples when studying sampling distributions?",
  answer("To generate outliers"),
  answer("To compare population and sample distributions"),
  answer("To see how much sample means vary from sample to sample", correct = TRUE),
  answer("To increase the sample size artificially"),
  correct = "The goal is to observe the variation across samples — that's the whole point of a sampling distribution.",
  allow_retry = TRUE
)

```
